\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-10 k-planar.tex
Suggested response to referees about 1-planarity.

\begin{thm}
\label{1-planar}
Every 1-planar graph is contained in $H\boxtimes P\boxtimes K_{30}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
\end{thm}

@@

Therefore each face of $G'$ is a $3$-cycle that contains at most one dummy vertex.  This implies that each face of $G$ is one of two types:
\begin{inparaenum}[(i)]
  \item bounded by three edges of $G$, none of which are crossed; or
  \item bounded by one edge of $G$ that is not crossed plus portions of two edges of $G$ that cross each other.
\end{inparaenum}

@@

By removing both spars from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three edges or $4$ edges; see \cref{one_planar_example}(b).  Observe that, for any spar $xy\in E(G)\setminus E(G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).  Therefore $\dist_{G'}(v,w)\le 2$.

@@

\referee{2}{Page 13.  Regarding the proof of Lemma 4, I think it is better to
include all the details from reference [14].  As far as I know, JCTB
does not have a page limit, so I do not see an issue with just
reproducing the entire proof and telling the reader that they can skip
all the details if they wish.  This is just my personal opinion
though, so the authors can ignore this request if they choose.}

\note{DW}{I agree we need more details. I am having trouble verifying this proof. }

>>>>

\note{PM}{Include in note to referee:  We choose to keep this section because
  $1$-planar graphs are of special interest to the graph drawing community.  For example, the annotated bibliography on $1$-planarity (\htmladdnormallink{https://arxiv.org/abs/1703.02261}{arXiv:1703.02261}) has 143 entries, and even this is now four years out of date.

  An important aspect of our result for $1$-planar graphs is that graph $H$ in the product is planar and has treewidth $3$, i.e., it has simple treewidth $3$.  In a more recent application of product structure \citet{BDJM} show that such graphs have $\ell$-vertex rankings using $O(\log n/\log\log\log n)$ colours, and this is tight. This result requires that $H$ have simple treewidth $3$, even treewidth $3$ is not sufficient.

  Finally, while cleaning up this proof we have also improved the constant from $30$ to $7$. The proof is quite similar to the previous proof except that graph $G'$ is obtained by removing both crossing edges involved in a crossing (the spars of a kite), rather than just one.  We thank the referee for forcing us to write this proof more carefully so that we could discover this improvement. 
}

\begin{thm}
\label{1-planar}
Every 1-planar graph is contained in $H\boxtimes P\boxtimes K_{7}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
\end{thm}

@@

Therefore each face of $G_0$ is a $3$-cycle that contains at most one dummy vertex.  This implies that each face of $G$ is one of two types:
\begin{inparaenum}[(i)]
  \item bounded by three edges of $G$, none of which are crossed; or
  \item bounded by one edge of $G$ that is not crossed plus portions of two edges of $G$ that cross each other.
\end{inparaenum}

@@

By removing both spars from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three or four edges; see \cref{one_planar_example}(b).  Observe that, for any spar $xy\in E(G)\setminus E(G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).  Therefore $\dist_{G'}(v,w)\le 2$.

@@

% \referee{2}{Page 13.  Regarding the proof of Lemma 4, I think it is better to
% include all the details from reference [14].  As far as I know, JCTB
% does not have a page limit, so I do not see an issue with just
% reproducing the entire proof and telling the reader that they can skip
% all the details if they wish.  This is just my personal opinion
% though, so the authors can ignore this request if they choose.}
% 
% \note{DW}{I agree we need more details. I am having trouble verifying this proof. }
